Tracking $p$-adic precision

### 2014-02-22 padicprec.tex
Proof of Proposition 3.9 improved

\newcommand{\Z}{\mathbb Z}
\newcommand{\Zp}{\Z_p}
\newcommand{\Q}{\mathbb Q}
\newcommand{\Qp}{\Q_p}
\newcommand{\Fp}{\mathbb{F}_p}
\newcommand{\R}{\mathbb R}
\renewcommand{\O}{\mathcal O}
\newcommand{\OK}{\mathcal{O}_K}
\newcommand{\XX}{\mathbf X}
\newcommand{\trans}{{}^{\text t}}
\newcommand{\T}{\mathcal{T}}

@@

\begin{proof}
Let $v \in \R$. We assume first that $\NP(f)^\star(v) = +\infty$. Then, 
using a multilinear version of Banach-Steinhaus Theorem, we can prove 
that $\sum_{n \geq 0} f_n$ does not converge on $B_E^-(e^v)$ and 
therefore that $\Lambda(f)(v) = +\infty$.

We now suppose that $\NP(f)^\star(v)$ is finite. To begin with, we 
assume in addition that $v$ is not a slope of $\NP(f)$. We set $u = 
\NP(f)^*(v)$. The function $m \mapsto \NP(f)(m) - vm + u$ has the 
following properties: 
\begin{enumerate}
\item it is piecewise affine and everywhere nonnegative,
\item it does not admit $0$ as a slope and
\item it vanishes at $x = n$ for some integer $n$ and $u = vn + \log \Vert f_n \Vert$.
\end{enumerate}
We deduce from these facts that there exists $c>0$ such that 
\[
vm-u \leq - \log \lVert f_m \rVert - c \cdot \vert n-m \vert
\]
for any $m\geq 0$.
Since $vm-u=vm-vn-\log \lVert f_n \rVert$, we get 
\[
-vn-\log \lVert f_n \rVert + c \cdot \vert n-m \vert \leq -vm-\log \lVert f_m \rVert.
\]
Therefore, for any $x \in  B_E(e^v)$ and $m \geq 0,$ we have
\[
\Vert f_m (x) \Vert \leq e^{-c \cdot \vert n-m \vert } \cdot
\lVert f_n \rVert \cdot e^{vn}< \lVert f_n \rVert \cdot e^{vn}.
\]
Thus, the series $\sum_{m \geq 0} f_m(x)$ converges and $\Vert f(x) 
\Vert \leq \Vert f_n \Vert \cdot e^{vn}$. We then get:
\begin{equation}
\label{eq:lowerLambda}
\Lambda (f)(v) \leq \log \big( \Vert f_n \Vert e^{vn} \big)
= vn + \log \Vert f_n \Vert = u.
\end{equation}
On the other hand, it follows from the definition of $\lVert f_n \rVert$ 
that there exists a sequence $(x_i)_{i\geq 0}$ in $B^-_E(e^v)$ such that 
$\lim_{i \to \infty} \Vert f_n (x_i) \Vert= \lVert f_n \rVert \cdot e^{vn}$. Since 
furthermore $\Vert f_m (x_i) \Vert \leq e^{-c \cdot \vert n-m \vert} 
\Vert f_n \Vert \cdot e^{vn}$ for all $i$, we get $\Vert f_m (x_i) \Vert 
< \Vert f_n (x_i) \Vert$ for $i$ large enough. For these $i$, we then 
have $\Vert f(x_i) \Vert = \Vert f_n (x_i) \Vert$. Passing to the limit 
on $i$, we find $\Lambda (f)(v) \geq u$. Comparing with 
\eqref{eq:lowerLambda}, we get $\Lambda (f)(v) = u = 
\NP(f)^\star(v)$.

>>>>

\newcommand{\N}{\mathbb N}
\newcommand{\Z}{\mathbb Z}
\newcommand{\Zp}{\Z_p}
\newcommand{\Q}{\mathbb Q}
\newcommand{\Qp}{\Q_p}
\newcommand{\Fp}{\mathbb{F}_p}
\newcommand{\R}{\mathbb R}
\renewcommand{\O}{\mathcal O}
\newcommand{\OK}{\mathcal{O}_K}
\newcommand{\XX}{\mathbf X}
\newcommand{\trans}{{}^{\text t}}
\newcommand{\T}{\mathcal{T}}

@@

\begin{proof}
Let $v \in \R$. We assume first that $\NP(f)^\star(v)$ is finite and 
that $v$ is not a slope of $\NP(f)$. We set $u = \NP(f)^*(v)$. The 
function $m \mapsto \NP(f)(m) - vm + u$ has the following properties:
\begin{enumerate}
\item it is piecewise affine and everywhere nonnegative,
\item it does not admit $0$ as a slope and
\item it vanishes at $x = n$ for some integer $n$ and $u = vn + \log \Vert f_n \Vert$.
\end{enumerate}
We deduce from these facts that there exists $c>0$ such that 
\[
vm-u \leq - \log \lVert f_m \rVert - c \cdot \vert n-m \vert
\]
for any $m\geq 0$.
Since $vm-u=vm-vn-\log \lVert f_n \rVert$, we get 
\[
-vn-\log \lVert f_n \rVert + c \cdot \vert n-m \vert \leq -vm-\log \lVert f_m \rVert.
\]
Therefore, for any $x \in  B_E(e^v)$ and $m \geq 0,$ we have
\[
\Vert f_m (x) \Vert \leq e^{-c \cdot \vert n-m \vert } \cdot
\lVert f_n \rVert \cdot e^{vn}< \lVert f_n \rVert \cdot e^{vn}.
\]
Thus, the series $\sum_{m \geq 0} f_m(x)$ converges and $\Vert f(x) 
\Vert \leq \Vert f_n \Vert \cdot e^{vn}$. We then get:
\begin{equation}
\label{eq:lowerLambda}
\Lambda (f)(v) \leq \log \big( \Vert f_n \Vert e^{vn} \big)
= vn + \log \Vert f_n \Vert = u.
\end{equation}
On the other hand, it follows from the definition of $\lVert f_n \rVert$ 
and the fact that $\vert K^\star \vert$ is dense in $\R$ (which holds
because $K$ is algebraically closed)
that there exists a sequence $(x_i)_{i\geq 0}$ in $B^-_E(e^v)$ such that 
$\lim_{i \to \infty} \Vert f_n (x_i) \Vert= \lVert f_n \rVert \cdot e^{vn}$. Since 
furthermore $\Vert f_m (x_i) \Vert \leq e^{-c \cdot \vert n-m \vert} 
\Vert f_n \Vert \cdot e^{vn}$ for all $i$, we get $\Vert f_m (x_i) \Vert 
< \Vert f_n (x_i) \Vert$ for $i$ large enough. For these $i$, we then 
have $\Vert f(x_i) \Vert = \Vert f_n (x_i) \Vert$. Passing to the limit 
on $i$, we find $\Lambda (f)(v) \geq u$. Comparing with 
\eqref{eq:lowerLambda}, we get $\Lambda (f)(v) = u = 
\NP(f)^\star(v)$.

\smallskip

We now assume that $\NP(f)^\star(v) = +\infty$. The function $x \mapsto 
\NP(f)(x) - vx$ is then not bounded from below. Since it is convex, it 
goes to $-\infty$ when $x$ goes to $+\infty$. Going back to the 
definition of $\NP(f)$, this implies that $v n + \log \Vert f_n \Vert$
goes to infinity when $n$ grows. It is then enough to establish the
following claim:
\begin{equation}
\label{eq:claim}
\forall n \in \N, \quad 
\Lambda(f)(v) \geq v n + \log \Vert f_n \Vert - \log 2.
\end{equation}
Let $n$ be a fixed integer. If $\Vert f_n 
\Vert = 0$, there is nothing to prove. Otherwise, let us consider an
element $x_n \in B^-_E(e^v)$ such that $\Vert f_n(x_n) \Vert \geq \frac 
1 2 \Vert f_n \Vert \cdot e^{vn}$. If the series $\sum_{m \geq 0} 
f_m(x_n)$ diverges, then $\Lambda(f)(v) = +\infty$ by definition and 
Eq.~\eqref{eq:claim} holds. On the other hand, if it converges, the
sequence $\Vert f_m(x_n) \Vert$ goes to $0$ when $m$ goes to infinity.
Hence it takes its maximum value $R$ a finite number of times; let us 
denote by $I \subset \N$ the set of the corresponding indices. For any
$\lambda \in B_K(1)$, the series defining $f(\lambda x_n)$ converges 
and
$$f(\lambda x_n) \in B_F(R)
\quad \text{and} \quad
f(\lambda x_n) \equiv \sum_{m \in I} \lambda^m f_m(x_n)
\pmod {B^-_F(R)}.$$
The quotient $B_F(R) / B^-_F(R)$ is a vector space over the residue
field of $K$, that is the field $k$ defined by $k = B_K(1)/B^-_K(1)$.
Hensel's lemma implies that $k$ is algebraically closed (because $K$ 
is) and then infinite. Therefore, there must exist $\lambda \in 
B_K(1)$ such that $\sum_{m \in I} \lambda^m f_m(x_n)$ does not vanish
in $B_F(R) / B^-_F(R)$. For such an element $\lambda$, we have $\Vert 
f(\lambda x_n) \Vert = R \geq \frac 1 2 \Vert f_n \Vert \cdot e^{vn}$. 
The claim \eqref{eq:claim} follows.

\smallskip
